---
title: "Check-in 3"
author: "Jack Rellamas"
date: "03-06-24"
format: 
  html:
    embed-resources: true
    code-tools: true
    code-summary: "Code"
---

Remember, **follow the instructions below and use R Markdown to create a pdf document with your code and answers to the following questions on Gradescope.** You may find a template file by clicking "Code" in the top right corner of this page.

Use the following code to download the annual consumer price index data from the World Bank:

```{r}
library(openxlsx)
cpi <- read.xlsx("https://thedocs.worldbank.org/en/doc/1ad246272dbbc437c74323719506aa0c-0350012021/original/Inflation-data.xlsx",
                 sheet = 5)
cpi <- cpi[1:203,]
```

1.  Create a line plot with Year on the x-axis and CPI on the y-axis, with different color lines for the US, Canada, and Mexico.

```{r}
library(tidyverse)
# make df filtering for USMCA countries
cpi_usmca <- filter(cpi, Country == "United States" |
                      Country == "Mexico" |
                      Country == "Canada")
ggplot(data = cpi_usmca,
       aes(x = "1970")) +
  geom_line()

```
