---
title: "Check-in 2"
author: "Jack Rellamas"
date: "02-28-24"
format: 
  html:
    embed-resources: true
    code-tools: true
    code-summary: "Code"
---

Remember, **follow the instructions below and use R Markdown to create a pdf document with your code and answers to the following questions on Gradescope.** You may find a template file by clicking "Code" in the top right corner of this page.

### Collaborators

INSERT NAMES OF ANY COLLABORATORS

## A. Flights Data

**1.** For this lab, we'll be using the `flights` data. You can load this data using the following code:

```{r}
library(nycflights13)
data(flights)
library(tidyverse)
```

**2.** Use the `mutate()` function to turn `origin` into a factor.

```{r}
flights <- mutate(flights, origin = factor(origin))
```

**3.** Compute the mean `arr_delay` for each origin airport. Which airport has the longest delays on average? You can drop missing rows.

```{r}
t <- flights %>% 
  group_by(origin) %>% 
  summarize(mean_arr_delay_origin = mean(arr_delay, na.rm = TRUE)) %>% 
  ungroup()
t
```

+----------+---------------------------+
| **\      | **mean_arr_delay_origin** |
| origin** |                           |
+:=========+:==========================+
|          |                           |
+----------+---------------------------+

|       |     |          |
|:------|:----|:---------|
| **1** | EWR | 9.107055 |
| **2** | JFK | 5.551481 |
| **3** | LGA | 5.783488 |

**4.** Use the `mutate()` function to turn `month` into a factor.

```{r}
flights <- mutate(flights, month = factor(month))
```

**5.** Compute the mean `arr_delay` by month. What do you notice?

```{r}
t <- flights %>% 
  group_by(month) %>% 
  summarize(mean_arr_delay_month = mean(arr_delay, na.rm = TRUE)) %>% 
  ungroup()
t
```

+---------+--------------------------+
| **\     | **mean_arr_delay_month** |
| month** |                          |
+:========+:=========================+
|         |                          |
+---------+--------------------------+

|        |     |            |
|:-------|:----|:-----------|
| **1**  | 1   | 6.1299720  |
| **2**  | 2   | 5.6130194  |
| **3**  | 3   | 5.8075765  |
| **4**  | 4   | 11.1760630 |
| **5**  | 5   | 3.5215088  |
| **6**  | 6   | 16.4813296 |
| **7**  | 7   | 16.7113067 |
| **8**  | 8   | 6.0406524  |
| **9**  | 9   | -4.0183636 |
| **10** | 10  | -0.1670627 |
| **11** | 11  | 0.4613474  |
| **12** | 12  | 14.8703553 |

**6.** Compute the mean `arr_delay` by month AND `origin`. What do you notice?

```{r}
t <- flights %>% 
  group_by(month, origin) %>% 
  summarize(mean_arr_delay_month_and_origin = mean(arr_delay, na.rm = TRUE)) %>% 
  ungroup()
t
```

EWR is usually has more delays than LGA and JFK.

All 3 of the airports experience a major increase in traffic and delays in June and July.

The best time to travel, if you want to avoid delays, is during the fall.

EWR has a significantly higher mean delay time in January than the other 2 airports.
